### 2024-03-25 doc/sections/mav-semantics.tex
[ tex ] more semantics

\begin{remark}
  In the case when the two pomonoids share a common unit (as in MAV),
  the last three conditions for a duoidal relationship are
  automatically satisfied.
\end{remark}

\subsection{MAV Algebras}

\begin{definition}
  An \emph{MAV Algebra} is a structure
  $(A, \leq, \otimes, ;, I, \lnot)$ with the following properties:
  \begin{enumerate}
  \item $(A, \leq, \otimes, I, \lnot)$ is $*$-autonomous and satisfies \emph{mix}.
  \item $(A, \leq, ;, I)$ is a pomonoid.
  \item $(\otimes, I)$ and $(;, I)$ are in a duoidal relationship.
  \item $(A, \leq)$ has binary meets, which we write as $x \with y$.
  \end{enumerate}
\end{definition}

@@

\begin{definition}\label{defn:mav-interpretation}
  Let $\mathit{At}$ be a set of atomic propositions. Given an MAV
  Algebra $(A, \leq, \otimes, \rhd, I, \lnot)$ and valuation
  $V : \mathit{At} \to A$, define the interpretation of MAV Formulas as follows:
  \begin{mathpar}
    FIXME
  \end{mathpar}
\end{definition}

@@

\subsection{MAV Frames}

To prove completeness of the cut-free fragment of SMAV, we will
construct a particular MAV Algebra from the syntax and cut-free
proofs. Since cut-free proofs do not a priori have all the necessary
structure for an MAV Algebra, we develop a procedure to construct MAV
Algebra from the lighter requirements of an MAV Frame, defined in this
section. MAV Frames have an intuitive reading as a kind of process
algebra, so we use process algebra-like notation. In the following
sections, we show how every MAV Frame induces an MAV Algebra, and in
Section XXX we will show that the MAV Algebra constructed from the
cut-free proof MAV Frame allows us to prove that all proofs in SMAV
have cut-free normal forms.

\begin{definition}
  An \emph{MAV Frame} is a structure
  $(F, \leq, I, \parallel, \rhd, +)$ where $(F, \leq)$ is a partial
  order, $(F, \leq, I, \parallel)$ is a commutative pomonoid,
  $(F, \leq, I, \rhd)$ is a pomonoid, $+$ is a binary monotone
  function on $(F, \leq)$, and these structures satisfy the following
  inequalities:
  \begin{enumerate}
  \item $(w \rhd x) \parallel (y \rhd z) \leq (w \parallel y) \rhd (x \parallel z)$
  \item $(x + y) \parallel z \leq (x \parallel z) + (y \parallel z)$
  \item $(w \rhd x) + (y \rhd z) \leq (w + y) \rhd (x + z)$
  \item $I + I \leq I$
  \end{enumerate}
\end{definition}

\begin{remark}
  \bob{A frame is nearly two duoidal structures, except that we don't
    assume that $+$ is a monoid, nor is there a unit element for it.}
\end{remark}

@@

\begin{proposition}
  \bob{The Cut-free system forms an MAV Frame; fwd ref to next section}
\end{proposition}

\subsection{Constructing MAV Algebras from MAV Frames}

\newcommand{\LowerSet}[1]{\widehat{#1}}

@@

\newcommand{\Day}[1]{\mathop{\widehat{#1}}}

@@

\begin{remark}
  When $(A, \leq)$ is an MAV Frame, by \ref{prop:day-construction} we
  obtain two pomonoids $(\Day{\parallel}, \Day{I})$ and
  $(\Day{\rhd}, \Day{I})$ on $\LowerSet{A}$. Moreover, the next
  proposition states that the duoidal relationship between these
  monoids is preserved by the Day construction:
\end{remark}

@@

The Lower Set adds residuals to the pomonoids of an MAV Frame, so the
Chu construction below will allow us to add duals to interpret all of
the multiplicative subsystem of MAV. To incorporate the additives in
such a way that we will be able to derive the proof normalisation
property, we need to refine lower sets to ones that are closed under
$+$-combinations.

@@

\begin{proposition}\label{prop:closed-monoid-distrib}
  Monoid construction 1, for monoids that distribute over $+$.
\end{proposition}

@@

To construct suitable MAV Algebras, we use the poset version of the
Chu construction \cite{barr}. The Chu construction builds
$*$-autonomous categories from symmetric monoidal closed categories
with pullbacks. In the preorder case, the requirement for preorders
simplifies to binary meets. Therefore, for this section, we let
$(A, \leq, \bullet, i, \rightblackspoon)$ be a symmetric monoidal
closed preorder with all binary meets.

@@

\begin{theorem}
  If $(F, \leq, \parallel, \rhd, I, +)$ is an MAV Frame, then
  $\Chu(\ClosedLowerSet{F}, I)$ has the structure of an MAV Algebra.
\end{theorem}

>>>>

\begin{remark}
  In the case when the two pomonoids share a common unit the last
  three conditions for a duoidal relationship are automatically
  satisfied.
\end{remark}

\bob{need to define \emph{entropy} and \emph{entropic} somewhere}

\subsection{MAV-algebras}

\begin{definition}
  An \emph{MAV-algebra} is a structure
  $(A, \leq, \otimes, ;, I, \lnot)$ with the following properties:
  \begin{enumerate}
  \item $(A, \leq, \otimes, I, \lnot)$ is $*$-autonomous and satisfies \emph{mix}.
  \item $(A, \leq, ;, I)$ is a pomonoid.
  \item $(\otimes, I)$ and $(;, I)$ are in a duoidal relationship.
  \item $(A, \leq)$ has binary meets, which we write as $x \with y$.
  \end{enumerate}
\end{definition}

@@

\begin{definition}\label{defn:mav-interpretation}
  Let $\mathit{At}$ be a set of atomic propositions. Given an
  MAV-algebra $(A, \leq, \otimes, \rhd, I, \lnot)$ and valuation
  $V : \mathit{At} \to A$, define the interpretation of MAV Formulas
  as follows:
  \begin{mathpar}
    FIXME
  \end{mathpar}
\end{definition}

@@

\subsection{MAV-frames}

To prove completeness of the cut-free fragment of SMAV, we will
construct a particular MAV-algebra from the syntax and cut-free
proofs. Since cut-free proofs do not a priori have all the necessary
structure for an MAV-algebra, we develop a procedure to construct
MAV-algebra from the lighter requirements of an MAV-frame, defined in
this section. MAV-frames have an intuitive reading as a kind of
process algebra, so we use process algebra-like notation. In the
following sections, we show how every MAV-frame induces an
MAV-algebra, and in Section XXX we will show that the MAV-algebra
constructed from the cut-free proof MAV-frame allows us to prove that
all proofs in SMAV have cut-free normal forms.

\begin{definition}
  An \emph{MAV-frame} is a structure
  $(F, \leq, I, \parallel, \rhd, +)$ where $(F, \leq)$ is a partial
  order, $(F, \leq, I, \parallel)$ is a commutative pomonoid,
  $(F, \leq, I, \rhd)$ is a pomonoid, $+$ is a binary monotone
  function on $(F, \leq)$, and these structures satisfy the following
  inequalities:
  \begin{enumerate}
  \item $(w \rhd x) \parallel (y \rhd z) \leq (w \parallel y) \rhd (x \parallel z)$
  \item $(x + y) \parallel z \leq (x \parallel z) + (y \parallel z)$
  \item $(w \rhd x) + (y \rhd z) \leq (w + y) \rhd (x + z)$
  \item $I + I \leq I$
  \end{enumerate}
\end{definition}

\begin{remark}
  \bob{A frame is nearly two duoidal structures, except that we don't
    assume that $+$ is a monoid, nor is there a unit element for
    it. If we assumed that $+$ was a lattice join, then (iii) and (iv)
    would automatically be true. We do not include this because these
    facts are not true for the cut-free MAV system.}
\end{remark}

@@

\begin{proposition}
  \bob{The Cut-free system forms an MAV-frame; fwd ref to next section}
\end{proposition}

\subsection{Constructing MAV-algebras from MAV-frames}

\newcommand{\LowerSet}[1]{\widehat{#1}}
\newcommand{\Day}[1]{\mathop{\widehat{#1}}}

@@

@@

\begin{remark}
  When $(A, \leq)$ is an MAV-frame, by \ref{prop:day-construction} we
  obtain two pomonoids $(\Day{\parallel}, \Day{I})$ and
  $(\Day{\rhd}, \Day{I})$ on $\LowerSet{A}$. Moreover, the next
  proposition states that the duoidal relationship between these
  monoids is preserved by the Day construction:
\end{remark}

@@

The Lower Set adds residuals to the pomonoids of an MAV-frame, so the
Chu construction below will allow us to add duals to interpret all of
the multiplicative subsystem of MAV. To incorporate the additives in
such a way that we will be able to derive the proof normalisation
property, we need to refine lower sets to ones that are closed under
$+$-combinations.

@@

\begin{proposition}\label{prop:closed-monoid-distrib}
  \bob{Monoid construction 1, for monoids that distribute over
    $+$. This also has a residual.}
\end{proposition}

@@

To construct suitable MAV-algebras, we use the poset version of the
Chu construction \cite{barr}. The Chu construction builds
$*$-autonomous categories from symmetric monoidal closed categories
with pullbacks. In the preorder case, the requirement for preorders
simplifies to binary meets. Therefore, for this section, we let
$(A, \leq, \bullet, i, \rightblackspoon)$ be a symmetric monoidal
closed preorder with all binary meets.

@@

\begin{theorem}
  If $(F, \leq, \parallel, \rhd, I, +)$ is an MAV-frame, then
  $\Chu(\ClosedLowerSet{F}, I)$ has the structure of an MAV-algebra.
\end{theorem}
